---
title: "had.27.46a20: Change in advice"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
fig_caption: yes
---

```{r, settings, include=FALSE,echo=FALSE}
rm(list=ls())

knitr::opts_chunk$set(
  fig.width = 9, message=FALSE, warning=FALSE,eval=TRUE,include=TRUE,echo=FALSE
)

options(digits = 2)
options(scipen=5)

library(icesTAF)
#library(stockassessment)
library(ggplot2)
library(RColorBrewer)
library(tidyverse)
#library(FLCore)
#library(icesAdvice)
#library(FLfse)
#library(icesSAG)
library(cowplot)
library(knitr)
library(kableExtra)

load("data/init.RData")
#ay <- 2025
advice_year <- ay+1
ages <- 0:8

#fig params
WIDTH <- 6
HEIGHT <- 5
UNITS <- "in"
MAR <- c(2,3.5,2.5,0.5)
MGP <- c(2,0.5,0)
PS <- 12
RESO <- 400

sourceDir("boot/software/utilities/")
output.dir <- "output/Change_in_advice/"

data_yrs <- 1972:(ay-1)

## Forecast parameters:
Ay <- (ay-3):(ay-1) # for biols
Sy <- (ay-3):(ay-1)  # for sel
Ry <- 2000:(ay-1) # for rec

ac<-as.character

source.now <- paste0("WGNSSK ",ay)
source.prev <- paste0("WGNSSK ",ay-1)

```


A comparison of the values used in this year's and last year's advice forecasts, for:
  
  * Forecast assumptions of SSB, Fbar, recruitment, total catch
* Stock weights-at-age used in the forecast years
* Selectivities used in the forecast
* Stock numbers-at-age, specifically for the forecast years (plot) and for all assessment and forecast years (table)
* Stock biomass-at-age, specifically for the forecast years (plot) and for all assessment and forecast years (table)


```{r, data,results='hide'}

# get Forecast assumptions
asmp.tab <- read.csv(paste0(output.dir,"Forecast_assumptions.csv"))

fWts <- read.csv(file = paste0(output.dir,"Forecast_stockwts.csv"))
fSel <- read.csv(file = paste0(output.dir,"Forecast_selectivity.csv"))

nAge <- read.csv(file=paste0(output.dir,"Forecast_N_at_age.csv"))
bAge <- read.csv(file=paste0(output.dir,"Forecast_B_at_age.csv"))

n_now <- read.csv(file=paste0(output.dir,"Now_assessment_N_at_age.csv"))
n_prev <- read.csv(file=paste0(output.dir,"Prev_assessment_N_at_age.csv"))

b_now <- read.csv(file=paste0(output.dir,"Now_assessment_B_at_age.csv"))
b_prev <- read.csv(file=paste0(output.dir,"Prev_assessment_B_at_age.csv"))

```

# Forecast assumptions for weight-at-age and selectivity
```{r, wts}

# compare forecast weights

p1 <- ggplot(data=fWts, aes(x=Age, y=Weight,colour=Source,group=Source)) + 
  facet_wrap(~Year,nrow = 2)+ geom_point()+
  geom_line() + theme_bw()+ labs(colour="",x="",y="mean weight (kg)") +
  theme(axis.title=element_text(size=8),axis.text=element_text(size=8),
        legend.text=element_text(size=9)) 
print(p1)

```

```{r,sel}
#compare selectivity

p1 <- ggplot(data=fSel, aes(x=Age, y=Selectivity,colour=Source,group=Source)) + 
geom_point()+
  geom_line() + theme_bw()+ labs(colour="",x="",y="Selectivity") +
  theme(axis.title=element_text(size=8),axis.text=element_text(size=8),
        legend.text=element_text(size=9)) 
print(p1)


```

# Change in numbers-at-age
```{r, n-at-age}

nAge$Type <- factor(nAge$Type,levels=c("Data","Intermediate year","Advice year"))
#compare n-at-age

p1 <- ggplot(nAge,aes(x=Age,y=N,group=interaction(Type,WG),colour=WG,shape=Type))+ geom_line()+geom_point(size=3)+labs(colour="",y="Abundance (thousands)",shape="")+
  facet_wrap(~Year,nrow=2)+theme_bw()+scale_shape_manual(values=c(16, 2, 0))

print(p1)


# n at age change table
dat.now <- n_now
dat.prev <- n_prev
colnames(dat.prev) <- colnames(dat.now) <- c("Year",ages[1:9])

# Combine with forecast N at age
# Forecast values ovewrite the assessment output here to account for when forecast values (e.g. median recruitment) are different to the assessment output
dat <- nAge
tmp.prev <- dat[dat$WG %in% source.prev,]
tmp.prev <- pivot_wider(tmp.prev,id_cols="Year",names_from="Age",values_from="N")
dat.prev <- bind_rows(dat.prev[dat.prev$Year < min(tmp.prev$Year),],tmp.prev)

tmp.now <- dat[dat$WG %in% source.now,]
tmp.now <- pivot_wider(tmp.now,id_cols="Year",names_from="Age",values_from="N")
dat.now <- bind_rows(dat.now[dat.now$Year < min(tmp.now$Year),],tmp.now)

# Find ratio
comp.yrs <- intersect(dat.now$Year,dat.prev$Year)
rat.n <- dat.now[dat.now$Year %in% comp.yrs,] / dat.prev[dat.prev$Year %in% comp.yrs,]
rat.n$Year <- comp.yrs

dat2tab <- rat.n[,as.character(ages)]
row.names(dat2tab) <- rat.n$Year

col.pal <- brewer.pal(n = 7, name = "RdBu")
# 0.8, 0.9, 0.95, 1, 1.05, 1.1, 1.2


colour_cols <- function(x){
  for (i in 1:dim(dat2tab)[2]){
    x <- x %>% column_spec(column=i+1, background = ifelse(dat2tab[,i] <= 0.80, col.pal[1],
                                                           ifelse(dat2tab[,i]<= 0.90,col.pal[2],
                                                                  ifelse(dat2tab[,i] <= 0.95,col.pal[3], 
                                                                         ifelse(dat2tab[,i] <= 1.05,col.pal[4],
                                                                                ifelse(dat2tab[,i] <= 1.1, col.pal[5],
                                                                                       ifelse(dat2tab[,i] <= 1.2,col.pal[6],col.pal[7])))))))
  }
  return(x)
}


dat2tab %>%
  kbl(longtable=T, booktabs = TRUE,caption = paste0("Ratio of stock numbers-at-age between forecasts (this year/lst year). Red colours indicate a decrease in numbers while blue colours indicate an increase")) %>%
  kable_styling(full_width=T,latex_options = c("repeat_header"),font_size = 12, repeat_header_continued = T, position = "center") %>%
  colour_cols()


```

# Change in biomass-at-age
```{r, biomass-at-age}

bAge$Type <- factor(bAge$Type,levels=c("Data","Intermediate year","Advice year"))
#compare b-at-age
bAge$Type <- factor(bAge$Type,levels=c("Data","Intermediate year","Advice year"))

# plot
ggplot(bAge,aes(x=Age,y=B,group=interaction(Type,WG),colour=WG,shape=Type))+ 
  geom_line() + geom_point(size=3) + labs(colour="",y="Biomass (tonnes)",shape="")+
  facet_wrap(~Year,nrow=2)+theme_bw()+scale_shape_manual(values=c(16, 2, 0))  

# Make a table to show the change in Biomass at age between forecasts as a ratio

# Read in B at age tables - as output from assessment (N * stock weights)
dat.prev <- b_prev
dat.now <- b_now

colnames(dat.prev) <- colnames(dat.now) <- c("Year",ages)

# Combine with forecast B at age
# Forecast values overwrite the assessment output here to account for when forecast values (e.g. median recruitment) are different to the assessment output
dat <- bAge
tmp.prev <- dat[dat$WG %in% source.prev,]
tmp.prev <- pivot_wider(tmp.prev,id_cols="Year",names_from="Age",values_from="B")
dat.prev <- bind_rows(dat.prev[dat.prev$Year < min(tmp.prev$Year),],tmp.prev)

tmp.now <- dat[dat$WG %in% source.now,]
tmp.now <- pivot_wider(tmp.now,id_cols="Year",names_from="Age",values_from="B")
dat.now <- bind_rows(dat.now[dat.now$Year < min(tmp.now$Year),],tmp.now)

# Find ratio
comp.yrs <- intersect(dat.now$Year,dat.prev$Year)
rat.b <- dat.now[dat.now$Year %in% comp.yrs,] / dat.prev[dat.prev$Year %in% comp.yrs,]
rat.b$Year <- comp.yrs

# print table
dat2tab <- rat.b[,as.character(ages)]
row.names(dat2tab) <- rat.b$Year

col.pal <- brewer.pal(n = 7, name = "RdBu")
# 0.8, 0.9, 0.95, 1, 1.05, 1.1, 1.2

colour_cols <- function(x){
  for (i in 1:dim(dat2tab)[2]){
    x <- x %>% column_spec(column=i+1, background = ifelse(dat2tab[,i] <= 0.80, col.pal[1],
                                                           ifelse(dat2tab[,i]<= 0.90,col.pal[2],
                                                                  ifelse(dat2tab[,i] <= 0.95,col.pal[3], 
                                                                         ifelse(dat2tab[,i] <= 1.05,col.pal[4],
                                                                                ifelse(dat2tab[,i] <= 1.1, col.pal[5],
                                                                                       ifelse(dat2tab[,i] <= 1.2,col.pal[6],col.pal[7])))))))
  }
  return(x)
}


dat2tab %>%
  kbl(longtable=T, booktabs = TRUE,caption = paste0("Ratio of stock biomass-at-age between forecasts (this year/lst year). Red colours indicate a decrease in biomass while blue colours indicate an increase")) %>%
  kable_styling(full_width=T, latex_options = c("repeat_header"),font_size = 12, repeat_header_continued = T, position = "center") %>%
  colour_cols()


```

# Compare forecast assumptions
```{r, assumptions}

# Compare forecast assumptions ------------------------------------------------------------

# read in and format forcast assumptions
dat <- asmp.tab
dat <- dat[dat$Year >(ay-2),]
dat$Type <- factor(dat$Type,levels=c("Data year","Intermediate year","Advice year"))
dat$Variable <- factor(dat$Variable,levels=c("SSB","Fbar","Total catch","Recruitment"))

# plot
ggplot(dat,aes(x=Year,y=Value,colour=WG,shape=Type))+geom_point(size=3)+facet_wrap(~Variable,scales="free_y")+
  theme_bw()+labs(x="",y="",colour="",shape="")+ scale_shape_manual(values=c(16, 2, 0))+ylim(0,NA)

```

